Sardine example: end to end truth for California Current

### 2019-05-22 FullSardineTruthEx.Rmd
add full sardine example

---
title: "Testing atlantisom: end to end truth for California Current"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

@@

@@

@@

@@

atBtxt2tidy <- atBtxt2 %>%
  select(Time:DIN) %>%
  #select(Time, FPL:DIN) %>%
  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
  gather(species, biomass, -Time) %>%
  filter(species %in% levels(surveyB_frombio$species))

@@

compareB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free") 
compareB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free") 
compareB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free") 
compareB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")

@@

@@

@@

@@

@@

## Test fishery functions: compare census with truth

Now for the fishery outputs. We will need total catch in tons for each species, and catch composition data (age and length). First we will test whether the catchbio output in `truth` matches the atlantis output file output[scenario.name]Catch.txt:

```{r testcatchbio1}

#just try the three species for now
catchbio_census_ss <- create_fishery_subset(dat = truth$catchbio,
                                         time = timeall,
                                         species = spp.name,
                                         boxes = boxall)

catchbio_census_ss_agg <- aggregate(atoutput ~ species + time,
                                    data=catchbio_census_ss, sum)

# learned the hard way this can be different from ecosystem outputs
fstepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutfinc

# what does the output look like?
catchB_outstep <- ggplot() +
  geom_line(data=catchbio_census_ss_agg, aes(x=time/fstepperyr,y=atoutput, color="catch atlantisom"),
            alpha = 10/10) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

catchB_outstep + 
  facet_wrap(~species, scales="free")

It doesnt. A hack gets it closer, suggesting that the problem is a bug in the output for models using codebases prior to late 2015 (CCA is mid-2015). From Beth Fulton's note in the wiki:

>Note that in CATCH.NC the numbers at age had issues in old versions. For versions pre-dating December 2015 you need to divide the numbers at age by (86400 * number_water_column_layers_in_the_box). New versions post December 2015 are now in numbers without needing further adjustment.

```{r testcatchbio2}

# should it be aggregated? I just did this above
catchbio_series <- sample_survey_numbers(catchbio_census_ss, surv_cv)

# if it is the pre-2015 bug mentioned in atlantis wiki, this may get it close
catchbio_series_adjust <- catchbio_series %>%
  mutate(atoutput = atoutput/(86400*4))

#does it match catch.txt (which should be in tons)?
atCtxt <- read.table(file.path(d.name, catch.file), header=T)
atCtxt <- atCtxt[, -grep("TsAct", colnames(atCtxt))] #hardcoded for catch.txt output

fishedlookup <- funct.groups %>%
  filter(IsFished > 0)

atCtxttidy <- atCtxt %>%
  rename_(.dots=with(fishedlookup, setNames(as.list(as.character(Code)), Name))) %>%
  gather(species, catchbio, -Time) %>%
  filter(species %in% levels(catchbio_series$species))

compare_catchB <-ggplot() +
  geom_line(data=catchbio_series_adjust, aes(x=time,y=atoutput, color="catch atlantisom"), 
            alpha = 10/10) +
  geom_point(data=atCtxttidy, aes(x=Time/365,y=catchbio, color="txt output catch bio"),
             alpha = 1/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

compare_catchB + 
  facet_wrap(~species, scales="free")

Test this with NOBA model, which is in a newer codebase that doesn't have this problem with the CATCH.nc file.

Update: NOBA has correct catch at age but the derived total catch biomass [still doesnt match catch.txt](https://sgaichas.github.io/poseidon-dev/TestCatchNOBA.html). I believe this is because we cannot multiply the correct weight at the layer the fishery catch is taken, which clearly does not match the median structn and resn across all model layers (as applied in `calc_biomass-age`. This may be buried deeply in the atlantis code, initialization, and prm files, but for now we will simply use the total catch weight out of atlantis catch.txt output.

## Catch output correction--apply to all species

Moving on to correcting catch at age for CCA:

```{r correctcatage}

# new atlantisom function to apply catage correction, or put inside run_truth

# first check that codebase is prior to corrected--to implement in atlantisom
#   read code version or date from log file
#   check against update date (Dec 12, 2015)
# grep("Atlantis SVN Last Change Date", readLines(file.path(d.name, "log.txt")), value = TRUE)

# alternative: less overhead if the system does it? log file is huge
# logfile <- paste0(file.path(d.name, "log.txt"))
# codedate <- system(paste0("grep 'Atlantis SVN' ", logfile))
# compare date in codedate in an if statement--fix in function

# read catch in nums output from truth$catch, or
# read in CATCH.nc (this from run_truth, modified with config files)

nc_catch <- paste0("output", scenario.name, 'CATCH.nc')
bboxes <- get_boundary(boxpars)

catch <- load_nc(dir = d.name,
                 file_nc = nc_catch,
                 bps = boxpars, #defined in census_spec.R
                 fgs = funct.groups,
                 select_groups = survspp,
                 select_variable = "Catch",
                 check_acronyms = TRUE,
                 bboxes = bboxes)

# find number of layers in each box
# The number of water column layers per box is in the initial conditions NC file:
# DIVinit_CalCurrentV3_Biol.nc : numlayers =

# read in initial conditions NC file
# this is already done for the other variables, modify from existing
# line 152 of load_nc
# num_layers <- RNetCDF::var.get.nc(ncfile = at_out, variable = "numlayers")[, 1]

# initial conditions file has different structure, so drop [,1]
at_init <- RNetCDF::open.nc(con = file.path(d.name, initial.conditions.file))

# Get info from netcdf file! (Filestructure and all variable names)
  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_init)$nvars - 1),
    function(x) RNetCDF::var.inq.nc(at_init, x)$name)
  numlayers <- RNetCDF::var.get.nc(ncfile = at_init, variable = "numlayers")

  RNetCDF::close.nc(at_init)

  # are these in box order??? if so make a box-numlayer lookup
  layerbox.lookup <- data.frame(polygon=boxall, numlayers)

  catch.tmp <- merge(catch, layerbox.lookup)

# divide the numbers at age by (86400 * number_water_column_layers_in_the_box)
# replace truth$catch atoutput with correction 
  catch <- catch.tmp %>%
    mutate(atoutput = atoutput / 86400 * numlayers) %>%
    select(species, agecl, polygon, time, atoutput)

  truth$catch <- catch

#save for later use, ar (better) resave truth file
#saveRDS(catch, file.path(d.name, paste0(scenario.name, "catchnums_corrected.rds")))
  save(truth,
      file = file.path(d.name, paste0("output", scenario.name, "run_truth.RData")))
```

Did the correction work? Not sure what to compare this to. First put it back through `calc_biomass_age` even though this won't match either?

```{r correctcatage-tobio, eval=FALSE}

biol <- load_biolprm(dir = d.name, file_biolprm = biol.prm.file)

# call within run_truth
catchbio <- calc_biomass_age(nums = catch,
    resn = truth$resn, structn = truth$structn, biolprm = biol)

#save for later use, even if wrong? takes a long time to generate
saveRDS(catchbio, file.path(d.name, paste0(scenario.name, "catchbio_ccatage.rds")))

It still doesn't match, similar to NOBA, but in the right ballpark now so the error is the median structn and resn called in the calc function above. This suggests that the catch at age in numbers is now correct, however.

```{r plotcorrected-bio}

catchbio <- readRDS(file.path(d.name, paste0(scenario.name, "catchbio_ccatage.rds")))

catchbio_corrected_agg <- aggregate(atoutput ~ species + time,
                                    data=catchbio, sum)

compare_catchB <-ggplot() +
  geom_line(data=catchbio_corrected_agg, aes(x=time,y=atoutput, color="catch atlantisom"), 
            alpha = 10/10) +
  geom_point(data=atCtxttidy, aes(x=Time/365,y=catchbio, color="txt output catch bio"),
             alpha = 1/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

compare_catchB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
compare_catchB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
compare_catchB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
compare_catchB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")

@@

Now we can sample fishery catch for lengths, ages, weights:

@@

This is true catch at age(class) for hake:

```{r catage2}

catchage <- catch_numssshigh %>%
  filter(species == "Mesopel_M_Fish")

Hake will need `calc_stage2age` applied because it has 2 true ages per age class.

@@

Hake catch lengths:

```{r vis-fishery-lf-test-2}

catchlen <- catch_length_census_ss$natlength %>%
  filter(species == "Mesopel_M_Fish")

>>>>

---
title: "Sardine example: end to end truth for California Current"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

@@

Catch in numbers has been corrected in the saved and reloaded truth file here. Ignore the catchbio component of this, it is not correct.

@@

Import the true catch biomass from a separate file. True total catch weight is currently available only on an annual and full-model (all polygons) basis:

```{r loadcatchb}

truecatchbio <- load_catch(d.name, file_catch = catch.file, fgs = funct.groups)

@@

surveyB_frombio_sardine <- surveyB_frombio[surveyB_frombio$species == "Pacific_sardine",]

@@

atBtxt2tidy <- atBtxt2 %>%
  select(Time:DIN) %>%
  #select(Time, FPL:DIN) %>%
  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
  gather(species, biomass, -Time) %>%
  filter(species %in% levels(surveyB_frombio_sardine$species))

@@

compareB + 
  facet_wrap(~species, scales="free")

@@

```{r survNtrue-plot}

surveyN <- readRDS(file.path(d.name, paste0(scenario.name, "surveyNcensus.rds")))

surveyN_sardine <- surveyN[surveyN$species == "Pacific_sardine",]

plotN <-ggplot() +
  geom_line(data=surveyN_sardine, aes(x=time/5,y=atoutput, color="survey census N"), 
            alpha = 10/10) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

plotN +
  facet_wrap(~species, scales="free")

@@

```{r Natage-plot}

Natage <- readRDS(file.path(d.name, paste0(scenario.name, "Natageclcensus.rds")))

Natage_sardine <- Natage[Natage$species == "Pacific_sardine",]

Natageplot <- ggplot(Natage_sardine, aes(x=agecl, y=atoutput)) +
  geom_point() +
  theme_tufte() +
  labs(subtitle = paste(scenario.name,
                        Natage_sardine$species))

Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")

@@

```{r lenatage-plot}

len <- length_census_ss$natlength %>%
  filter(species == "Pacific_sardine")

lfplot <- ggplot(len, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte() +
  labs(subtitle = paste(scenario.name,
                        len$species))

@@

#length_census_ss <- readRDS(file.path(d.name, paste0(scenario.name, "length_census_sardhakeboca.rds")))

@@

## Fishery catch

We found [after much testing](https://sgaichas.github.io/poseidon-dev/Atlantisom2SSworkflowtest.html) that the most reliable source for catch in biomass at present is the Atlantis output catch.txt file, loaded above. Here is sardine annual total catch:

```{r plotcatch}

truecatchbio_sardine <- truecatchbio[truecatchbio$species == "Pacific_sardine",]

plotcatch <- ggplot() +
  geom_line(data=truecatchbio_sardine, aes(x=time,y=atoutput, color="true catch bio"),
            alpha = 10/10) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

plotcatch +
  facet_wrap(~species, scales="free")

@@

Now (that catch nums are corrected!) we can sample fishery catch for lengths, ages, weights:

@@

@@
